### 2024-10-10 LLMs/manuscript/main.tex
first results

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}

\title{Zero-shot LLMs can save additional work in machine learning prioritised screening for systematic reviews}
\author[1,2]{Max Callaghan}

\affil[1]{Mercator Research Institute on Global Commons and Climate Change, Torgauer Straße 12-15, 10829 Berlin, Germany}
\affil[2]{%
	Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, 14412 Potsdam, Germany
}

\begin{document}
	\maketitle

	\doublespacing

	\begin{abstract}
		Zero shot learning using LLMs has received widespread attention in recent years, including in the automation of systematic review screening. 
		However, there has been insufficient research on robust processes to automate screening responsibly. 
		For systematic review screening, responsible automation requires that the risks of missing relevant studies (which are inherent when the process is automated) can be accurately quantified and managed.
		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use robust stopping criteria to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
		We find that that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can save on average [x] percentage points of additional work compared to a baseline model.
		Across the whole database of [x] systematic review datasets, which would have required [x] hours to screen manually, LLM-assisted screening can reduce the time required to [x] hours, with the next best model reducing the time required to [y] hours, while still ensuring a 95\% chance of missing no more than 5\% of relevant studies.
	\end{abstract}

	\section*{Introduction}

	Systematic Reviews are important tools to synthesise findings across multiple studies, enabling decision-makers to make decisions that are informed by the best available evidence. 
	A key part of the process of systematic reviews involves finding studies using a transparent process that aims to be as comprehensive as possible. 
	Usually, bibliographic records are searched for using structured queries across multiple databases \cite{lefebvre_chapter_2023}.
	This results in a list of records that may or not be relevant to the review. The titles and abstracts of these records are then examined by hand - often by two independent reviewers - and it is judged whether they should be included, which means they are read again at the full text level, or excluded.
	This is a laborious and repetitive process - known as screening - and in times of increasing numbers of publications, this becomes more and more challenging.

	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of this process, in order to reduce the labour required to complete a systematic review. However, much of this literature demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}.

	This is reflected in the fact that the guidelines for conducting systematic reviews published by the Cochrane colloquium \cite{lefebvre_chapter_2023} do not recommend that machine learning is used to save labour during screening. This is because it remains uncertain how the use of automation may result in relevant studies being missed.

	Early attempts to automate systematic review screening used simple models like logistic regression and support vector machines. Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs). However, not enough attention has been paid to comparing LLM-based approaches with the much simpler (and computationally less expensive) alternatives. Further, papers which have investigated the use of LLMs have so far not considered how they might be used in a responsible manner where evaluation data is not available.

	For example, in \cite{wang_zero-shot_2024}, the authors introduce a method to generate probability-like scores for the inclusion of each document, by retrieving the probabilities of the next token after prompting the model. The prompt gives information on the review and provides the title and abstract of the record in question, and asks the model to answer whether it should be included with a ``yes'' or a ``no''. The probability that the next token is ``no'' is subtracted from the probability that the next token is ``yes'', resulting in the inclusion score. This process has two advantages over simply prompting the model and recording whether it says yes or no. First, the result becomes deterministic, in that a single model will always give the same score for the same prompt. Second, in returning a probability-like score instead of a binary response, we are better able to discriminate between documents considered highly likely to be irrelevant and those considered mildly likely to be irrelevant by the model.

	However, the paper proposes a strategy to save work using the model that is not able to accurately quantify or mitigate the risk of missing studies. Documents with an inclusion score greater than a threshold paramater $\theta$ are screened by hand, while all remaining documents are automatically excluded. The methods proposed to set the threshold paramater (either extrapolating from other reviews, or by setting the paramater such that it achieves the desired recall level on a set of predefined relevant documents) are \textit{not} able to reliably ensure that a desired level of recall is met.

	Although the LLM-based procedure is compared to a fine-tuned BERT-like model, the details of this are not clear, and it appears that there is no comparison to the most common mode of machine learning assisted screening, known as machine learning prioritised screening. With ML-prioritised screening, a subset of documents are screened by hand, and these documents are used to train a model to predict the relevance of remaining unscreened documents. A batch of those records with the highest predictions are then screened by hand, before the model is retrained using the expanded training dataset. This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.

	In this paper, we adapt the probabilistic zero-shot screening process using LLMs to a ML-prioritised screening setting, and show that - when combined with statistically valid stopping criterion - it is able to reliably achieve high levels of recall while saving work. Secondly, we compare LLM-assisted approaches to simpler methods available, both in terms of the work saved, and in terms of computational requirements.

	\section*{Methods}

	\subsection*{Probabilistic inclusion scores from LLMs}

	Following \cite{wang_zero-shot_2024}, we write a prompt template (see supplementary material) that ``asks'' a large language model whether a document $d$ should be included in a review $t$. Instead of using the prompt to generate a response, we pass the prompt to the model and extract the next token probabilities. The inclusion score $S$ for document $d$ in review $t$ is given by subtracting the probability that the next token is ``no'' $P(no|d,t)$ from the probability that the next token is ``yes'' $P(yes|d,t)$.

	\begin{equation}S(d,t) = P(yes|d,t) - P(no|d,t)\end{equation}

	In a new review, one could generate scores in this way for each document returned by a query, and screen these in descending order of their inclusion score, applying a stopping criterion to decide when remaining unscreened documents can be excluded automatically.  We simulate this process using the Synergy dataset of completed reviews \cite{de_bruin_synergy_2023}, and report on the work that would have been saved, and the actual level of recall that would have been achieved, given a range of recall targets and uncertainty levels.

	We use the stopping criteria in \cite{callaghan_statistical_2020}\footnote{Using the python package available at \url{https://buscarpy.readthedocs.io/en/latest/}}, which uses the hypergeometric distribution to test a null hypothesis that a target level of recall has \textit{not} been achieved. When the 1 minus the p score returned by the criterion is below the our confidence level, which indicates that it would have been unlikely to observe the previous sequence of inclusion and exclusion decisions had we missed our recall target, screening is stopped.

	\subsection*{Active learning baselines}

	With the LLM-assisted screening approach, predictions are made once for each document, and are made solely on the basis of the prompt. With active learning approaches, which are more commonly deployed, predictions are made using a model based on records that have already been screened, and the model is frequently retrained as more screening data becomes available, meaning that predictions are generated multiple times for many documents.

	We implement our active learning pipeline as follows for each model and each review dataset, where each dataset has $N$ unique documents. Screening here, in a simulation context, refers to revealing the screening decision that was made when the dataset was created.

	\begin{enumerate}
		\item ``Screen'' a sample of 10\% of papers in the review dataset. If no papers in the sample are relevant, keep screening until at least one relevant paper has been screened.
		\item Train the model using the screened documents and prediction inclusion for remaining unscreened documents
		\item Screen the $N/10$ documents with the highest predicted scores 
		\item Repeat steps 2 and 3 until all documents have been screened
	\end{enumerate}

	This process generates an ordered list of documents that corresponds to the order in which documents would have been seen if our approach was being applied in the real world. Using this ordered list, we calculate stopping points for a variety of recall targets and confidence levels.

	Because the random sample can affect the quality of the predictions and the resulting rankings, we repeat this experiment 100 times, each with a different random seed, for each model and dataset combination.

	\section*{Results}

	\begin{figure}
		\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}
		\caption{Rankings produced by LLM-assisted screening (coloured lines) and by active learning using a support vector machine. Left panel shows the recall achieved at different proportions of relevant documents screened. Right panel shows the distribution of ROC AUC scores for the support vector machine approach (boxplot) and for LLMs (coloured crosses)}
		\label{fig:rankings}
	\end{figure}

	First, we start by examining the rankings produced by our ML-assisted screening processes. Figure \ref{fig:rankings} shows these for a selected review dataset on Psychological theories of depressive relapse and recurrence, which contained 38,114 records, of which 0.2\% were relevant. Looking at the left side, we see that all evaluated models result in the identification of all relevant documents before all documents have been screened, meaning that work savings are in principle possible. On the right-hand side, we see the Receiver Operating Characteristic - Area Under the Curve (ROC-AUC) score, which measures the likelihood that a random relevant document is ranked higher than a random irrelevant record. A score of 1 indicates that all relevant documents are correctly identified first, before any irrelevant documents have been seen. A score of 0.5 indicates that the ranking is no better than chance.

	Overall, across the X datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}). Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer paramaters. This difference was more pronounced for Llama2 than for Llama3.1

	\begin{figure}
		\includegraphics[width=\columnwidth]{../../figures/macro_comparison.pdf}
		\caption{The distribution of ROC AUC scores across datasets. The results for the support vector machine (SVM) display the median score for each dataset.}
		\label{fig:macro}
	\end{figure}

	\section*{Discussion}

	\bibliographystyle{unsrt}

	\bibliography{../stopping-criteria}
	%\bibliography{../LLMs}

	\appendix

	\section*{Supplementary Material}

	\subsection*{Prompt template}

\end{document}

### 2024-10-10 LLMs/poster/poster_GES_2024.tex
first results

% Created with Brian Amberg's LaTeX Poster Template. Please refer for the     %
% attached README.md file for the details how to compile with `pdflatex`.     %
% --------------------------------------------------------------------------- %
% $LastChangedDate:: 2011-09-11 10:57:12 +0200 (V, 11 szept. 2011)          $ %
% $LastChangedRevision:: 128                                                $ %
% $LastChangedBy:: rlegendi                                                 $ %
% $Id:: poster.tex 128 2011-09-11 08:57:12Z rlegendi                        $ %
% --------------------------------------------------------------------------- %
\documentclass[a0paper,portrait]{baposter}

\usepackage{relsize}		% For \smaller
\usepackage{url}			% For \url
\usepackage{epstopdf}	% Included EPS files automatically converted to PDF to include with pdflatex
\usepackage{moresize}
\usepackage{multicol}
\usepackage[font=small,labelfont=bf]{subcaption} % <-- changed to subcaption
\usepackage{natbib}

\renewcommand{\bibsection}{}

%%% Global Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\graphicspath{{pix/}}	% Root directory of the pictures 
\tracingstats=2			% Enabled LaTeX logging with conditionals

%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{bordercol}{RGB}{40,40,40}
\definecolor{headercol1}{HTML}{DEEEB2}
\definecolor{headercol2}{RGB}{80,80,80}
\definecolor{headerfontcol}{RGB}{0,0,0}
\definecolor{boxcolor}{HTML}{ffffff}

%\definecolor{orange}{HTML}{FF7F00}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Utility functions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Save space in lists. Use this after the opening of the list %%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
	\setlength{\itemsep}{1pt}
	\setlength{\parskip}{0pt}
	\setlength{\parsep}{0pt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Document Start %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
	\hyphenation{systematic}
\typeout{Poster rendering started}

%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\background{
%	\begin{tikzpicture}[remember picture,overlay]%
%	\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%	{\includegraphics[height=1.1\textheight]{background}};
%	\end{tikzpicture}
}

%%% General Poster Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% Eye Catcher, Title, Authors and University Images %%%%%%%%%%%%%%%%%%%%%%
\begin{poster}{
%	grid=false,
	columns=3,
	colspacing=4.2mm,
	headerheight=0.08\textheight,
	background=none,
	eyecatcher=false,
	%posterbox options
	headerborder=closed,
	borderColor=black,
	headershape=rectangle,
	headershade=plain,
	headerColorOne=blue,
	textborder=rectangle,
	boxshade=plain,
	boxColorOne=white,
	headerFontColor=white,
	headerfont=\color{white}\large\bfseries\sffamily,
	textfont=\larger\sffamily,
	linewidth=1pt
}
%%% Eye Cacther %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
	Eye Catcher, empty if option eyecatcher=false - unused
}
%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\sf\bf
	\vspace{1em}The safe use of LLMs for screening in systematic reviews
}
%%% Authors %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
 %\url{https://dx.doi.org/10.1038/s41558-019-0684-5}\\

}
%%% Logo %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
% The logos are compressed a bit into a simple box to make them smaller on the result
\setlength\fboxsep{0pt}
\setlength\fboxrule{0pt}
	\fbox{
		\begin{minipage}[border]{14em}
			\centering
			\includegraphics[width=6.5em, ]{../pres/images/MCC_Logo_RZ_rgb.jpg}
		\end{minipage}
	}
}

\headerbox{The promise of LLMs}{name=1,column=0,row=0}{

%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}

LLMs have appeared useful for various tasks, but how can we \textbf{safely} use them to save work while screening?

\medskip

They are temperamental, and they return text, not probabilites
}

\headerbox{Next token probabilities}{name=2,column=0,row=0, below=1}{

	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}

	LLMs are \textbf{probabilistic models}, which calculate the \textbf{probable next token} given a \textbf{prompt}.

	\medskip

	\cite{wang_zero-shot_2024} showed that we can ask a yes/no question and get a probability-like score by subtracting the probability that the next token is no, from the probability it is yes.

	\[
	S(d,t) =
	P(\text{yes}|d,t) - P(\text{no}|d,t)
	\]

	\medskip

	But the paper did not contain a way to use these scores to save work in a way that would reliably satisfy our need in systematic reviews for high recall

}

\headerbox{Prioritised screening}{name=3,column=0,row=0, below=2}{

	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}

	However, if we have probabistic scores for documents, then we can simply do prioritised screening \textit{with} a reliable \textbf{stopping criteria} \citep{callaghan_statistical_2020}
	\smallskip

		\includegraphics[width=\linewidth]{../../figures/stopping.pdf}

}

\headerbox{Results}{name=21,column=1, span=2, row=0}{
	Using the synergy dataset \citep{de_bruin_synergy_2023}, we compared rankings from LLM screening with rankings generated in a traditional ``active learning'' pipeline with SVMs

	\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}

	\includegraphics[width=\linewidth]{../../figures/Jeyaraman_2020.pdf}

	The best models mostly, but not always, outperform the baseline
}

\headerbox{Results}{name=22,column=1, span=1, below=21}{

	Llama 3.1 works better than 2 \cite{wang_zero-shot_2024}, and larger models work better than smaller

	\includegraphics[width=\linewidth]{../../figures/macro_comparison.pdf}

}

\headerbox{Conclusion}{name=32,column=2, span=1, below=21}{

	Performance with 0 human input is impressive. Combining approaches, or using human labels to provide in-context learning could be promising.

}

\headerbox{Bibliography}{name=bib,column=0,row=0, below=3}{
	\scriptsize

	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}
	\bibliographystyle{apalike}
	\bibliography{../LLMs}
}

\begin{posterbox}[name=footerbox,span=2,column=0,above=bottom,boxheaderheight=0em,textborder=none]{}
{
	For more info on stopping criteria see \url{mcallaghan.github.io/buscar-app} \\
	Max Callaghan \\
{\smaller callaghan@mcc-berlin.net, @MaxCallaghan5}
}
\end{posterbox}

\end{poster}
\end{document}

### 2024-10-10 LLMs/poster/poster_amld_2020.tex
first results

--- 2024-10-10 LLMs/poster/poster_GES_2024.tex
@@
 \usepackage[font=small,labelfont=bf]{subcaption} % <-- changed to subcaption
-\usepackage{natbib}
-
-\renewcommand{\bibsection}{}
 
@@
 \begin{document}
-	\hyphenation{systematic}
 \typeout{Poster rendering started}
@@
 {\sf\bf
-	\vspace{1em}The safe use of LLMs for screening in systematic reviews
+	Biased Urns for efficient Stopping Criteria in technology-Assisted Reviews (BUSCAR)
 }
@@
 
-\headerbox{The promise of LLMs}{name=1,column=0,row=0}{
+\headerbox{Problem}{name=problem,column=0,row=0}{
 
@@
 
-LLMs have appeared useful for various tasks, but how can we \textbf{safely} use them to save work while screening?
+Machine-learning priorised screening can save work, but only if we know when it is \textbf{safe} to stop
 
-\medskip
+	\includegraphics[width=\linewidth]{../pres/images/potential_stopping.pdf}
 
-They are temperamental, and they return text, not probabilites
 }
 
-\headerbox{Next token probabilities}{name=2,column=0,row=0, below=1}{
+\headerbox{Urns}{name=urn,column=0,below=problem}{
 
@@
 
-	LLMs are \textbf{probabilistic models}, which calculate the \textbf{probable next token} given a \textbf{prompt}.
+	The hypergeometric distribution models the likelihood of retrieving $k$ red marbles from an urn containing $N$ marbles, of which $K$ are red, if we pick out $n$ marbles without replacing them.
 
-	\medskip
+	\includegraphics[width=\linewidth]{../pres/images/Keats_urn.jpg}
 
-	\cite{wang_zero-shot_2024} showed that we can ask a yes/no question and get a probability-like score by subtracting the probability that the next token is no, from the probability it is yes.
-
-	\[
-	S(d,t) =
-	P(\text{yes}|d,t) - P(\text{no}|d,t)
-	\]
-
-	\medskip
-
-	But the paper did not contain a way to use these scores to save work in a way that would reliably satisfy our need in systematic reviews for high recall
+	Thinking of relevant documents as red marbles, this provides a conservative stopping criterion (Callaghan and Müller-Hansen, 2020), when we assume (conservatively) that the likelihood of drawing a red or white marble is equal.
 
@@
 
-\headerbox{Prioritised screening}{name=3,column=0,row=0, below=2}{
+\headerbox{Biased Urns}{name=biasedurn,column=1,span=2,row=0}{
 
@@
 
-	However, if we have probabistic scores for documents, then we can simply do prioritised screening \textit{with} a reliable \textbf{stopping criteria} \citep{callaghan_statistical_2020}
-	\smallskip
+Biased Urn Theory (Fog, 2023), models what happens when the likelihood of drawing a random item of interest is higher than the likelihood of drawing a random irrelevant item.
+The ratio between these two is the \textbf{bias}
+}
 
-		\includegraphics[width=\linewidth]{../../figures/stopping.pdf}
+\headerbox{Method}{name=method,column=1, span=2, below=biasedurn}{
+
+	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}
+
+	\begin{enumerate}
+		\item Estimate maximum number of relevant documents in a given confidence interval from random sample
+		\item Look back at the last 1/2 of documents since the random sample, and use Wallenius' non-central hypergeometric distribution and Maximum Likelihood Estimation to find the level of bias below which 95\% of the probability distribution occurs
+		\item Look back at previous batches (last 1, last 2, last 3...) to see the likelihood of the observed number or relevant documents occuring given the null hypothesis that we have missed our target.
+	\end{enumerate}
+}
+
+\headerbox{Experiments}{name=exp,column=1, span=2,below=method}{
+
+	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}
+
+	We test our stopping criteria with 100 runs on complete systematic review datasets (from Cohen, 2006), with different recall targets (0.8, 0.95, 0.99) and different confidence levels (0.8, 0.95, 0.99).
+We show the distribution of additional work savings compared to an omniscient criterion, as well as the distribution of recall, both on average, and for each dataset.
+}
+
+\headerbox{Results}{name=res,column=1, span=2,below=exp}{
+
+%\setcaptiontype{figure}% Fake a figure environment
+%
+%	\subfloat[Target recall 0.8 - conf 0.8]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.8_0.8.pdf}}  
+%	\hfill
+%	\subfloat[Confidence 0.95]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.8_0.95.pdf}}
+%	\hfill
+%	\subfloat[Confidence 0.99]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.8_0.99.pdf}}
+%
+%	\subfloat[Target recall 0.95 - conf 0.8]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.95_0.8.pdf}}  
+%\hfill
+%\subfloat[Confidence 0.95]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.95_0.95.pdf}}
+%\hfill
+%\subfloat[Confidence 0.99]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.95_0.99.pdf}}
+%
+%	\subfloat[Target recall 0.99 - conf 0.8]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.99_0.8.pdf}}  
+%\hfill
+%\subfloat[Confidence 0.95]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.99_0.95.pdf}}
+%\hfill
+%\subfloat[Confidence 0.99]{
+%	\includegraphics[width=0.32\textwidth]{../plots/jp_conservative_bias_0.99_0.99.pdf}}	
+%
+%  %\caption{Signed distance field representation}
+
+\includegraphics{../plots/composite.pdf}
+
+Using a biased urn results in higher work savings (see example pathway, while still maintaining recall levels above a target level at a given confidence level).
 
@@
 
-\headerbox{Results}{name=21,column=1, span=2, row=0}{
-	Using the synergy dataset \citep{de_bruin_synergy_2023}, we compared rankings from LLM screening with rankings generated in a traditional ``active learning'' pipeline with SVMs
-
-	\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}
-
-	\includegraphics[width=\linewidth]{../../figures/Jeyaraman_2020.pdf}
-
-	The best models mostly, but not always, outperform the baseline
-}
-
-\headerbox{Results}{name=22,column=1, span=1, below=21}{
-
-	Llama 3.1 works better than 2 \cite{wang_zero-shot_2024}, and larger models work better than smaller
-
-	\includegraphics[width=\linewidth]{../../figures/macro_comparison.pdf}
-
-}
-
-\headerbox{Conclusion}{name=32,column=2, span=1, below=21}{
-
-	Performance with 0 human input is impressive. Combining approaches, or using human labels to provide in-context learning could be promising.
-
-}
-
-\headerbox{Bibliography}{name=bib,column=0,row=0, below=3}{
-	\scriptsize
-
-	%\includegraphics[width=\linewidth]{../plots/literature_size/pubs_time_wgb_lp.pdf}
-	\bibliographystyle{apalike}
-	\bibliography{../LLMs}
-}
-
-\begin{posterbox}[name=footerbox,span=2,column=0,above=bottom,boxheaderheight=0em,textborder=none]{}
+\begin{posterbox}[name=footer,span=2,column=0,above=bottom,boxheaderheight=0em,textborder=none]{}
 {
-	For more info on stopping criteria see \url{mcallaghan.github.io/buscar-app} \\
-	Max Callaghan \\
+	Email me for how to use the criterion in your project. R and python packages forthcoming. \\
+	Max Callaghan \& Finn Müller-Hansen\\
 {\smaller callaghan@mcc-berlin.net, @MaxCallaghan5}

### 2024-10-23 llm_screen.py
run experiments

@@ -1,0 +1,130 @@
+import typer
+import time
+from datetime import datetime
+import torch
+import pandas as pd
+from transformers import LlamaForCausalLM, AutoConfig, pipeline, BitsAndBytesConfig, AutoTokenizer
+from string import Template
+from synergy_dataset import iter_datasets
+import pyarrow.parquet as pq
+
+
+def binary_probs(tokenizer, model, prompt, no_words=['no'], yes_words=['yes'], return_all=False, print_answer=False):
+    device = 'cuda' if torch.cuda.is_available() else 'cpu'
+    encoded_text = tokenizer(prompt, return_tensors="pt").to(device)
+    #1. step to get the logits of the next token
+    with torch.inference_mode():
+        outputs = model(**encoded_text)
+
+    if print_answer:
+        out = model.generate(**encoded_text, max_new_tokens=1024)
+        print(tokenizer.decode(out[0]))   
+    
+    next_token_logits = outputs.logits[0, -1, :]
+    
+    # 2. step to convert the logits to probabilities
+    next_token_probs = torch.softmax(next_token_logits, -1)
+    
+    topk_next_tokens= torch.topk(next_token_probs, 50)
+    tokens = [tokenizer.decode(x).strip().lower() for x in topk_next_tokens.indices]
+    p = topk_next_tokens.values
+    
+    df = pd.DataFrame.from_dict({'t': tokens,'p': p.cpu()})
+    y = df[df['t'].isin(yes_words)]['p'].sum()
+    n = df[df['t'].isin(no_words)]['p'].sum()
+
+    if return_all:
+        return df.groupby('t').sum().reset_index().sort_values('p', ascending=False).reset_index(drop=True)
+    return y, n
+
+def generate(model, tokenizer, text):
+    t = tokenizer(text, return_tensors="pt").to('cuda')
+    out = model.generate(**t)
+    print(tokenizer.decode(out[0]))
+
+def main(model_name: str):
+    """
+    Run evaluations testing ML prioritisation with LLMs
+    """
+    quantization_config = BitsAndBytesConfig(
+        load_in_4_bit=True,
+        bnb_4bit_use_double_quant=True,
+        bnb_4bit_quant_type="nf4",
+        bnb_4bit_compute_dtype=torch.bfloat16
+    )
+    tokenizer = AutoTokenizer.from_pretrained(model_name)
+    #config = AutoConfig.from_pretrained(
+    #    model,
[…]

### 2024-10-23 mpi_test.py

@@ -1,0 +1,20 @@
+from mpi4py import MPI
+import pyarrow as pa
+import pyarrow.parquet as pq
+import numpy as np
+
+comm = MPI.COMM_WORLD
+rank = comm.Get_rank()
+print(rank)
+
+#f = h5py.File('p/tmp/maxcall/ml-screening/parallel_test.hdf5','w',driver='mpio',comm=MPI.COMM_WORLD)
+
+#dset = f.create_dataset('test', (4,), dtype='i')
+#dset[rank] = rank
+
+size = 10000
+x = np.zeros(size)
+x[:] = rank
+y = np.arange(size)
+table = pa.table({'rank': x, 'y': y})
+pq.write_to_dataset(table,root_path='/p/tmp/maxcall/ml-screening/ptest',partition_cols=['rank'])

### 2024-10-23 pipeline.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-
-#SBATCH --job-name=ml-screening-pipeline
-#SBATCH --qos=short
-#SBATCH --time=00:01:00
-#SBATCH --output=out/pipeline.out
-#SBATCH --error=out/pipeline.err
-
-job1=$(sbatch --parsable -J "screen-simulation" --ntasks $1 run_prioritisation.sh $1 "$2")
-#pred2=$(sbatch --parsable --dependency=afterok:$pred1 -J "$JNAME" run_predictions.sh $MODEL "$VAR")
-	

### 2024-10-23 setup_db.py

@@ -1,8 +1,11 @@
 import sqlite3
 import typer
+from synergy_dataset import Dataset, iter_datasets
+import json
 
 def main(delete: int=0):
     # Set up database
+    con = sqlite3.connect('/p/tmp/maxcall/ml-screening/experiments.sql')
     con = sqlite3.connect('output_data/experiments.sql')
     create_runs = '''
     CREATE TABLE IF NOT EXISTS runs (
@@ -16,18 +19,51 @@ def main(delete: int=0):
         id INTEGER PRIMARY KEY,
         rec_id TEXT,
         screened_order INTEGER,
-        review TEXT,
+        first_prediction FLOAT,
+        last_prediction FLOAT,
+        review_id INTEGER,
         relevant INTEGER,
-        run_id
+        run_id INTEGER
     );
+    '''
+    create_batch_preds = '''
+    CREATE TABLE IF NOT EXISTS batch_predictions (
+        id INTEGER PRIMARY KEY,
+        review_id INTEGER,
+        prediction FLOAT,
+        n_screened INTEGER,
+        run_id INTEGER
+    );
+    '''
+    create_reviews = '''
+    CREATE TABLE IF NOT EXISTS reviews (
+        review_id INTEGER PRIMARY KEY,
+        review_name TEXT,
+        n_records INTEGER,
+        prevalence FLOAT
+    )
     '''
     cur = con.cursor()
     if delete>0:
         print('dropping tables')
         cur.execute('DROP TABLE IF EXISTS runs')
         cur.execute('DROP TABLE IF EXISTS ordered_records')
+        cur.execute('DROP TABLE IF EXISTS batch_predictions')
+        cur.execute('DROP TABLE IF EXISTS reviews')
+        cur.execute("VACUUM")
+        con.commit()
     cur.execute(create_runs)
     cur.execute(create_or)
+    #cur.execute(create_batch_preds)
+    cur.execute(create_reviews)
+    res = cur.execute('PRAGMA journal_mode=WAL')
+    
+    with open('pipelines.json','r') as f:
+        pipelines = json.load(f)
+        print(pipelines)
+    
+
+    cur.close()
     return
 
 

### 2024-10-23 slurm/gpu_setup.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#SBATCH --qos=gpumedium                                                                           
+#SBATCH --time=00:10:00
+
+#SBATCH --output=out/setup.out
+#SBATCH --error=out/setup.err
+#SBATCH --ntasks=1
+#SBATCH --partition=gpu                      
+#SBATCH --gres=gpu:v100:1
+
+export I_MPI_PMI_LIBRARY=/p/system/slurm/lib/libpmi.so
+
+module load anaconda/2023.09
+conda create --name llm-screening python=3.11
+source activate llm-screening
+pip install -r llm_requirements.txt

### 2024-10-23 slurm/llm-pipeline.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+#SBATCH --job-name=ml-screening-pipeline-2-70b
+#SBATCH --qos=gpushort
+#SBATCH --partition=gpu
+#SBATCH --gres=gpu:2
+#SBATCH --time=23:50:00
+#SBATCH --output=out/pipeline-l2-70b.out
+#SBATCH --error=out/pipeline-l2-70b.err
+
+module load python/3.12.3
+
+source venv/bin/activate
+
+export HF_HOME=/p/tmp/maxcall/hf/
+
+# python llm_screen.py meta-llama/Meta-Llama-3.1-8B-Instruct
+# python llm_screen.py meta-llama/Meta-Llama-3.1-70B-Instruct
+# python llm_screen.py meta-llama/Llama-2-7b-chat-hf
+python llm_screen.py meta-llama/Llama-2-70b-chat-hf

### 2024-10-23 slurm/mpitest.sh

@@ -1,0 +1,8 @@
+module load python/3.12.3
+module load intel/oneAPI/2024.0.0
+#module load openmpi/5.0.3
+#module load hdf5/1.14.4
+
+source cpuvenv/bin/activate
+
+mpirun -np 100 python mpi_test.py

### 2024-10-23 slurm/pipeline.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+#SBATCH --job-name=ml-screening-pipeline
+#SBATCH --qos=short
+#SBATCH --time=00:01:00
+#SBATCH --output=out/pipeline.out
+#SBATCH --error=out/pipeline.err
+
+job1=$(sbatch --parsable -J "screen-simulation" --ntasks $1 slurm/run_prioritisation.sh $1 "$2")
+clean=$(sbatch --parsable --dependency=afterok:$job1 slurm/clean_parquet.sh)
+	

### 2024-10-23 stopping.py

@@ -1,0 +1,113 @@
+import pyarrow as pa
+import pyarrow.parquet as pq
+from buscarpy import retrospective_h0
+import time
+import typer
+import sqlite3
+import pandas as pd
+from itertools import product
+import numpy as np
+
+def main(version: str):
+    """
+    Run evaluations testing ML prioritisation across multiple ML pipelines and multiple datasets
+    """
+
+    try:
+        from mpi4py import MPI
+        comm = MPI.COMM_WORLD
+        rank = comm.Get_rank()
+        print(rank)
+        test = False
+    except:
+        import random
+        rank = 0
+        print(f'Test mode: randomly assigned {rank}')
+        test = True
+
+    db_path = 'output_data/experiments.sql'
+    con = sqlite3.connect(db_path)
+    rdf = pd.read_sql_query('SELECT * FROM reviews',con)
+
+    cur = con.cursor()
+    stmt = f'SELECT run_id FROM runs WHERE version="{version}" AND thread="{rank}";'
+    cur.execute(stmt)
+    res = cur.fetchone()
+    run_id = res[0]
+
+    batch_size = 10
+
+    for i, row in rdf.iterrows():
+
+        print(row['review_id'])
+    
+        table = pq.read_table(
+            '/p/tmp/maxcall/ml-screening/ordered_records', 
+            filters=[
+                ('review_id', '=', row['review_id']),
+                ('run_id', '=', run_id)
+            ]
+        ).to_pandas().sort_values('screened_order').reset_index(drop=True)
+        for recall_target in [0.8,0.9,0.95,0.99]:
+            pdf = pd.DataFrame(
+                retrospective_h0(
+                    table['relevant'], table.shape[0], 
+                    batch_size=batch_size, plot=False, recall_target=recall_target,
+                    confidence_level=0.99
+                )
+            ).rename(columns={'p': f'p_{recall_target}'})
+            if f'p_{recall_target}' in table.columns:
+                table = table.drop(columns=f'p_{recall_target}')
+            table = table.merge(
+                pdf, left_on='screened_order', right_on='batch_sizes', how='left'
+            ).drop(columns='batch_sizes').ffill()
+        table.to_parquet(
+            '/p/tmp/maxcall/ml-screening/ordered_records',
+            partition_cols=['review_id','run_id'],
+            existing_data_behavior='delete_matching'
+        )
+
+    # models = pq.read_table(
[…]

### 2024-10-23 LLMs/pres/LLM-screening.tex

\documentclass[9pt,aspectratio=169]{beamer}
\mode<presentation>
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{listings}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}

\usepackage{hyperref}

\usetheme{Singapore}
%\usecolortheme{seahorse}

\usefonttheme{professionalfonts}

\title[Chatbots Annotators]{LLM-prioritised screening}

\author{Max Callaghan}
\institute[MCC]{
\includegraphics[height=1cm,width=2cm]{images/MCC_Logo_RZ_rgb.jpg}
}

\newtheorem*{remark}{}

\bibliographystyle{apalike}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\section{Introduction}

\section{Implementation}

\begin{frame}{We can ask LLMs whether a study should be included in a systematic review}
	\begin{columns}
		\begin{column}{0.5\linewidth}
			\begin{figure}
				\footnotesize
				%\includegraphics[width=\linewidth]{images/example.txt}
				\input{images/example.txt}
			\end{figure}
		\end{column}
		\begin{column}{0.5\linewidth}
			\only<2->{\begin{figure}
					\includegraphics[width=\columnwidth]{images/example_probs.pdf}
			\end{figure}}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Zero-shot Generative Large Language Models for Systematic Review Screening Automation}

	\cite{wang_zero-shot_2024} propose a method to extract probability-like scores from LLMs for inclusion/exclusion decisions in a systematic review

	\begin{figure}
		\includegraphics[width=\columnwidth]{images/fig1}
	\end{figure}
\end{frame}

\begin{frame}{Results are mixed, because the probabilities are poorly calibrated, and their attempts to re-invent stopping criteria do not makes sense}

	\begin{figure}
		\includegraphics[width=0.8\columnwidth]{images/tab3-1}
	\end{figure}
\end{frame}

\begin{frame}[fragile]{Implementation is relatively straightforward}
\small
\begin{lstlisting}[language=Python]
def binary_probs(tokenizer, model, prompt, no_words=['no'], yes_words=['yes'], return_all=False):
    device = 'cuda' if torch.cuda.is_available() else 'cpu'
    encoded_text = tokenizer(prompt, return_tensors="pt").to(device)
    #1. step to get the logits of the next token
    with torch.inference_mode():
        outputs = model(**encoded_text)

    next_token_logits = outputs.logits[0, -1, :]

    # 2. step to convert the logits to probabilities
    next_token_probs = torch.softmax(next_token_logits, -1)

    topk_next_tokens= torch.topk(next_token_probs, 50)
    tokens = [tokenizer.decode(x).strip().lower() for x in topk_next_tokens.indices]
    p = topk_next_tokens.values

    df = pd.DataFrame.from_dict({'t': tokens,'p': p.cpu()})
    y = df[df['t'].isin(yes_words)]['p'].sum()
    n = df[df['t'].isin(no_words)]['p'].sum()

    if return_all:
        return df.groupby('t').sum().reset_index().sort_values('p', ascending=False).reset_index(drop=True)
    return y-n, y+n
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]{Implementation is relatively straightforward}
	\small
\begin{lstlisting}[language=Python]
prompt = Template('''<s>[INST] <<SYS>>
You are a systematic review helper tasked with finding out whether a study is relevant to the review $t

Answer 'yes' if the study is relevant, or 'no' if not
<</SYS>>

Study: $s

Should the study be included? Answer yes or no. [/INST] ''')

prompt.substitute({'t': review, 's': study_title}),
\end{lstlisting}

\end{frame}

\section{Results}

\begin{frame}{Result \#1: It works!}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> The method produces a ranking that identifies all relevant documents before all documents have been screened
				\item<2-> But it doesn't perform as well as simple tfidf + SVM
				\item<3-> To compare the quality of rankings, we can use ROC AUC scores
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\only<1>{\includegraphics[width=\linewidth]{../../figures/LLM_Brouwer_simple.pdf}}
			\only<2->{\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}}
			%\only<3>{}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{This method simply produces a ranking of documents, but is this ranking better than what we can produce with traditional active learning approaches?}

	\begin{columns}
		\begin{column}{0.618\linewidth}
			\begin{figure}

				\only<1>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019_p_distribution.pdf}}
				\only<2>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019.pdf}}
				\only<3>{\includegraphics[width=\columnwidth]{../../figures/Bos_2018.pdf}}
				\only<4>{\includegraphics[width=\columnwidth]{../../figures/Brouwer_2019.pdf}}
				\only<5>{\includegraphics[width=\columnwidth]{../../figures/Chou_2004.pdf}}
				\only<6>{\includegraphics[width=\columnwidth]{../../figures/Donners_2021.pdf}}
				\only<7>{\includegraphics[width=\columnwidth]{../../figures/Jeyaraman_2020.pdf}}
			\end{figure}
		\end{column}
		\begin{column}{0.382\linewidth}
			\begin{itemize}
				\item<1-> We generated LLM inclusion scores for each document in a set of reviews in the Synergy dataset
				\item<2-> We compare what happens to screening in descending order of this ranking to 100 active learning runs with a default configuration of SVM with Tfidf
				\item<3-> Initial results are not very promising
				\item<5->But LLMs can do well where traditional methods struggle
				\item<7->Sometimes they are no use at all
			\end{itemize}
		\end{column}
	\end{columns}

\end{frame}

\begin{frame}{All results}
	\begin{columns}
	\begin{column}{0.618\linewidth}
		\begin{figure}

			\only<1->{\includegraphics[width=\columnwidth]{../../figures/llm_svm_roc.pdf}}

		\end{figure}
	\end{column}
	\begin{column}{0.382\linewidth}
		\begin{itemize}
			\item<1-> LLMs and basic active learning pipelines seem to have different weakneses
			\item<2-> Combining both could improve general performance
			\item<3-> LLMs seem most useful for smaller datasets (where active learning has little time to learn)
		\end{itemize}
	\end{column}
\end{columns}
\end{frame}

\section{Conclusion}

\begin{frame}{Some things still to explore}

\begin{itemize}
	\item<1-> Prompting strategies (inclusion criteria)
	\item<2-> Bigger/different models
	\item<3-> Combining LLMs with traditional approaches
	\item<4-> Updating prompts based on user feedback
\end{itemize}

\end{frame}

\begin{frame}{Conclusions}
	\begin{columns}
		\begin{column}{0.4\linewidth}
			\begin{itemize}
				\item<1-> LLMs are neither a quick fix or a silver bullet
				\item<2-> Evaluation is vital
				\item<3-> We can't forget the need for appropriate stopping criteria \cite{callaghan_statistical_2020}
				\item<4->\url{mcallaghan.github.io/buscar-app}
			\end{itemize}
		\end{column}
		\begin{column}{0.6\linewidth}
			\begin{figure}
				\only<3>{\includegraphics[width=\columnwidth]{../../figures/stopping.pdf}}
			\end{figure}
		\end{column}
	\end{columns}

\end{frame}

\begin{frame}
	Thanks!

	\bigskip

	\hrule

	\bigskip

	\url{callaghan@mcc-berlin.net}

	\medskip

	Twitter: @MaxCallaghan5

	\bigskip

	\begin{figure}
	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
	\end{figure}
\end{frame}

\begin{frame}
	\bibliography{../LLMs}
\end{frame}

\end{document}

### 2024-10-23 LLMs/pres/LLM-screening_wed_seminar.tex

--- 2024-10-23 LLMs/pres/LLM-screening.tex
@@
 
-\title[Chatbots Annotators]{LLM-prioritised screening}
+\title[LLM screening and coding]{Screening and coding with LLMs}
 
@@
 
-\section{Introduction}
+\begin{frame}{Outlook}
+	\begin{itemize}
+		\item<1-> Results from the implementation of \cite{wang_zero-shot_2024}
+		\item<2-> Additional results from Santiago's thesis project on coding
+	\end{itemize}
+\end{frame}
 
-\section{Implementation}
+%\section{Introduction}
+%
+%\section{Implementation}
+
+\section{Screening}
 
@@
 
-\section{Results}
+%\section{Results}
 
@@
 
-\section{Conclusion}
+%\section{Conclusion}
 
@@
 				\item<3-> We can't forget the need for appropriate stopping criteria \cite{callaghan_statistical_2020}
-				\item<4->\url{mcallaghan.github.io/buscar-app}
+				%\item<4->\url{mcallaghan.github.io/buscar-app}
 			\end{itemize}
@@
 
-\begin{frame}
-	Thanks!
+\section{Coding}
 
-	\bigskip
+\begin{frame}{Using LLMs for coding}
+		\begin{columns}
+		\begin{column}{0.4\linewidth}
+			\begin{itemize}
+				\item<1-> Santiago's thesis showed us that LLMs can achieve comparable performance with BERT-type models
+				\item<2->They need no training data to achieve this performance
+				\item<3->But we do need annotated data for ``prompt engineering'' ($\approx$ training?), and for evaluation
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.6\linewidth}
+			\begin{figure}
+				\only<1->{\includegraphics[width=\columnwidth]{../../figures/LLM-coding.pdf}}
+			\end{figure}
+		\end{column}
+	\end{columns}
 
-	\hrule
+\end{frame}
 
-	\bigskip
+\begin{frame}{Using LLMs for coding}
+	\begin{columns}
+		\begin{column}{0.4\linewidth}
+			\begin{itemize}
+				\item<1-> Results were aggregated Macro F1 scores for 30 labels
+				\item<2-> Aggregate results hide much variation
+				\item<3-> This varies across the different coding categories
+				\item<4-> And there seems to be a clear correlation with number of labels
+			\end{itemize}
+		\end{column}
+		\begin{column}{0.6\linewidth}
+			\begin{figure}
+				\only<2>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs.pdf}}
+				\only<3>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs-categories.pdf}}
+				\only<4>{\includegraphics[width=\columnwidth]{../../../LLM-eval/gpt/figures/BERT-vs-LLMs-n.pdf}}
+			\end{figure}
+		\end{column}
+	\end{columns}
 
-	\url{callaghan@mcc-berlin.net}
+\end{frame}
 
-	\medskip
-
-	Twitter: @MaxCallaghan5
-
-	\bigskip
-
-	\begin{figure}
-	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
-	\end{figure}
+	\begin{frame}{Conclusions}
 \end{frame}
+%	
+%\begin{frame}
+%	Thanks!
+%	
+%	\bigskip
+%	
+%	\hrule
+%	
+%	\bigskip
+%	
+%	\url{callaghan@mcc-berlin.net}
+%	 
+%	\medskip
+%	
+%	Twitter: @MaxCallaghan5
+%	
+%	\bigskip
+%	
+%	\begin{figure}
+%	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
+%	\end{figure}
+%\end{frame}
+%
 

### 2024-10-24 LLMs/manuscript/main.tex
analysis of stopping criteria

@@
 \usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage[margin=2.5cm]{geometry}
 
-\title{Zero-shot LLMs can save additional work in machine learning prioritised screening for systematic reviews}
+\title{Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?}
 \author[1,2]{Max Callaghan}
@@
 
-	Overall, across the X datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}).
-Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer paramaters.
-This difference was more pronounced for Llama2 than for Llama3.1
-
 \begin{figure}
@@
 
+\begin{table}
+	\centering
+	\input{../../tables/recall.tex}
+	\caption{Recall achieved at the 1-confidence percentile across model runs. Values below the recall target are highlighted in bold.}
+	\label{tab:recall}
+\end{table}
+
+	Overall, across the X datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}).
+Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer paramaters.
+This difference was more pronounced for Llama2 than for Llama3.1.
+
+	Next, we apply the stopping criteria to all prioritised runs across all model and dataset combinations, and analyse the recall achieved and the work savings that would have been made.
+In table \ref{tab:recall}, we show the recall achieved at the percentile of model runs given by 1 minus the confidence interval, for each targeted recall level.
+	For example, if we had used SVM and had stopped screening with $p=0.5$ for a recall target of $0.8$, we would have achieved a recall level of at least $0.94$ 50\% of the time.
+In other words, we would have missed our recall target \textit{less than} 50\% of the time, meaning that our confidence estimation was conservative (in line with expectations from \cite{callaghan_statistical_2020}).
+	Moreover, across 100 combinations of recall target, confidence level, and model, the target was only missed more than would be expected given the confidence level in 3 cases (highlighted in bold text).
+Moreover, for each of these cases the percentiles are highly imprecise, as they are calculated over the distribution of 26 datasets.
+For the case of SVM, where there are 100 runs for each dataset, the target was never missed more than would have been expected given the confidence level.
+This highlights that the stopping criteria developed in \cite{callaghan_statistical_2020} provide a highly effective way to quantify and mitigate the risks of missing relevant studies.
+
+	\begin{table}
+		\centering
+		\input{../../tables/wss_p.tex}
+		\caption{Average proportional work savings across datasets for each combination of  model recall target and confidence level.
+For the SVM column where there are multiple runs per dataset, we first calculate the median value for each dataset before aggregating across datasets.}
+		\label{tab:wssp}
+	\end{table}
+
+	We now turn to the work savings that would have been achieved at the stopping points given by the stopping criteria.
+Table \ref{tab:wssp} shows the mean and standard deviation of the work savings achieved for each combination of model, recall target, and confidence level.
+For low recall targets and low levels of confidence, large work savings are possible (>50\%).
+Increasing either the recall target or the level of confidence reduces the work savings that can be achieved.
+For example, for a recall target of 99\%, it is still possible to achieve work savings of 28.8\%, as long as one accepts a 50-50 chance of missing the target.
+However, if one wants to be 99\% confident that a 99\% recall target has not been missed, on average one would only save 1.6\% of the work.
+
+	\begin{table}
+		\centering
+		\input{../../tables/wsst.tex}
+		\caption{Average proportional work savings across datasets for each combination of  model recall target and confidence level.
+For the SVM column where there are multiple runs per dataset, we first calculate the median value for each dataset before aggregating across datasets.}
+		\label{tab:wsst}
+	\end{table}
+
+	Despite the previous results showing that the largest and newest LLMs achieved superior ROC-AUC scores to a simpler SVM model with iterative retraining, once we look at the work savings that would have been achieved in a realistic setting, we see that in the vast majority of settings, we would still save more work with a simpler SVM model.
+For 17 of the 20 combinations of recall targets and confidence levels shown in table \ref{tab:wssp}, the highest work savings (highlighted in bold) are achieved with an SVM.
+
+	Finally, we look at total work savings available across datasets (table \ref{tab:wsst}). The Synergy dataset contains systematic review datasets that range from 258 to 48,375 records. 
+%	Because of the relationship between the number of records that have been seen and the confidence intervals calculated as part of the stopping criterion, larger proportional work savings are in principle possible with larger datasets \cite{callaghan_statistical_2020}.
+	Screening each of the datasets by hand would have required 169,288 records to be screened.	
+	If we had aimed for a recall target of 80\% and accepted a 50\% chance of missing the target, this number could have been reduced by 100,248 in the worst performing model (Llama 2 7b) to 132,808 in the best performing model (Llama 3.1 70b).
+	Again, increasingly ambitious recall targets and increasingly high levels of confidence reduce the amount of work that can be saved substantially.
+It is still possible to achieve an ambitious recall target of 90\% with 90\% confidence and save almost half of the otherwise required work (63,245 records).
+However, in highly ambitious settings, such as 99\% recall with 99\% confidence, only minimal work savings are possible (7,980 records, or just under 5\%)
+
 	\section*{Discussion}
+
+	These results add to previous evidence that it is possible to \textit{safely} use prioritised screening approaches that can guarantee a given level of recall with a given level of confidence \cite{callaghan_statistical_2020}.
+Across a wide range of datasets, as well as a wide range of different models, the stopping criteria ensured that the targeted level of recall was missed no more than expected given the confidence level.
+
+	Next, we add to promising evidence that LLMs can produce remarkably good rankings of abstracts for screening, even with very little human input.
+However, when we consider a practical use-case, the benefits of using LLMs for screening seem to be much more limited.
+In the majority of settings, higher work savings are achieved with much simpler SVM models.
+
+	Simplified version, no inclusion criteria - but also simplified SVM setting, no parameter tuning, large batches, no finetuned.
+
+	Possibility to combine
+
+	Scientific treatment of the costs and benefits of different options.
 

### 2024-11-01 LLMs/pres/2024_10_30_LLM-screening.tex
manuscript updates

--- 2024-10-23 LLMs/pres/LLM-screening.tex
@@
 \usepackage{listings}
+\usepackage{booktabs}
 
@@
 
-\title[Chatbots Annotators]{LLM-prioritised screening}
+\title[LLM screening]{Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?}
 
@@
 
+%\section{Introduction}
+%
+%\section{Implementation}
+
 \section{Introduction}
 
-\section{Implementation}
+\begin{frame}{Can LLMs help us to save more work?}
 
-\begin{frame}{We can ask LLMs whether a study should be included in a systematic review}
 \begin{columns}
 	\begin{column}{0.5\linewidth}
+		\begin{itemize}
+			\item<1->Prioritised screening has saved us a lot of work
+			\item<2->But we still spend substantial amounts of time screening documents for systematic reviews
+			\item<3->LLMs have been proposed as a solution, with
+			\begin{itemize}
+				\item<4-> \cite{wang_zero-shot_2024} showing impressive results for the zero-shot setting
+				\item<5-> and  \cite{xia_llmscreen_2024} erroneously claiming that LLMs fix all of the problems faced until now by ml+screening
+			\end{itemize} 
+			\item<6-> There are a lot of bad papers, which will continue to proliferate, each tweaking some parameters to achieve a high performance that cannot be generalised
+		\end{itemize}
+	\end{column}
+
+	\begin{column}{0.5\linewidth}
+		\only<4>{
 		\begin{figure}
-				\footnotesize
-				%\includegraphics[width=\linewidth]{images/example.txt}
-				\input{images/example.txt}
+			\includegraphics[width=\columnwidth]{images/tab3-1}
 		\end{figure}
-		\end{column}
-		\begin{column}{0.5\linewidth}
-			\only<2->{\begin{figure}
-					\includegraphics[width=\columnwidth]{images/example_probs.pdf}
-			\end{figure}}
+		}	
+		\only<5->{
+					\begin{figure}
+				\includegraphics[width=\columnwidth]{images/xia}
+			\end{figure}
+	}
 	\end{column}
 \end{columns}
+
 \end{frame}
+
+\begin{frame}{Realistic and useful evaluations of LLMs for screening}
+
+	Despite the hype, LLMs might still be helpful, but we need to know
+
+	\begin{itemize}
+		\item<2->How do LLMs compare against commonly used (and much cheaper!) baselines?
+		\item<3->How could they safely be used in the real world (Hint: with stopping criteria and prioritised screening)
+	\end{itemize}
+\end{frame}
+
+\section{Prioritisation with LLMs}
 
@@
 		\includegraphics[width=\columnwidth]{images/fig1}
-	\end{figure}
-\end{frame}
-
-\begin{frame}{Results are mixed, because the probabilities are poorly calibrated, and their attempts to re-invent stopping criteria do not makes sense}
-
-	\begin{figure}
-		\includegraphics[width=0.8\columnwidth]{images/tab3-1}
 	\end{figure}
@@
 \begin{frame}[fragile]{Implementation is relatively straightforward}
+
 	\small
@@
 
+\begin{frame}[fragile]{Implementation is relatively straightforward}
+
+	\begin{columns}
+	\begin{column}{0.5\linewidth}
+		\begin{figure}
+			\footnotesize
+			%\includegraphics[width=\linewidth]{images/example.txt}
+			\input{images/example.txt}
+		\end{figure}
+	\end{column}
+	\begin{column}{0.5\linewidth}
+		\only<2->{\begin{figure}
+				\includegraphics[width=\columnwidth]{images/example_probs.pdf}
+		\end{figure}}
+	\end{column}
+\end{columns}
+
+\end{frame}
+
+\section{Comparison}
+
+\begin{frame}{Comparing priorisation approaches}
+
+\begin{itemize}
+	\item<1-> We evaluate using the Synergy dataset \cite{de_bruin_synergy_2023} of 26 systematic reviews (Tim, Diana, Sergio, Lena, and James, will make a much bigger one!)
+	\item<2-> For each review, we generate 100 different initial random samples of 10\%, and do normal prioritised screening in batches of 10\% with an SVM 100 times
+	\item<3-> We also make a prediction for each document with 4 Llama models of different sizes and ages
+	\item<4-> We turn these predictions into a prioritised list of documents to be screened
+\end{itemize}
+
+\end{frame}
+
 \section{Results}
@@
 				\item<1-> The method produces a ranking that identifies all relevant documents before all documents have been screened
-				\item<2-> But it doesn't perform as well as simple tfidf + SVM
-				\item<3-> To compare the quality of rankings, we can use ROC AUC scores
+				\item<2-> The first model we tried was much worse than our SVM, but newer and bigger models were much more promising!
+				\item<3-> But performance varied across datasets
+				\item<4-> ROC AUC scores offer one way to compare rankings
 			\end{itemize}
@@
 			\only<1>{\includegraphics[width=\linewidth]{../../figures/LLM_Brouwer_simple.pdf}}
-			\only<2->{\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}}
+			\only<2>{\includegraphics[width=\linewidth]{../../figures/Brouwer_2019.pdf}}
+				\only<3>{\includegraphics[width=\columnwidth]{../../figures/Bos_2018.pdf}}
+				\only<4>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019.pdf}}
+			\only<5>{\includegraphics[width=\columnwidth]{../../figures/Chou_2004.pdf}}
+			\only<6>{\includegraphics[width=\columnwidth]{../../figures/Donners_2021.pdf}}
+			\only<7>{\includegraphics[width=\columnwidth]{../../figures/Jeyaraman_2020.pdf}}
+			\only<8>{\includegraphics[width=\columnwidth]{../../figures/macro_comparison.pdf}}
 			%\only<3>{}
@@
 
-\begin{frame}{This method simply produces a ranking of documents, but is this ranking better than what we can produce with traditional active learning approaches?}
+%\begin{frame}{All results}
+%	\begin{columns}
+%	\begin{column}{0.618\linewidth}
+%		\begin{figure}
+%			
+%			\only<1->{\includegraphics[width=\columnwidth]{../../figures/llm_svm_roc.pdf}}
+%
+%		\end{figure}
+%	\end{column}
+%	\begin{column}{0.382\linewidth}
+%		\begin{itemize}
+%			\item<1-> LLMs and basic active learning pipelines seem to have different weakneses
+%			\item<2-> Combining both could improve general performance
+%			\item<3-> LLMs seem most useful for smaller datasets (where active learning has little time to learn)
+%		\end{itemize}
+%	\end{column}
+%\end{columns}
+%\end{frame}
 
+\begin{frame}{Stopping criteria}
 	\begin{columns}
-		\begin{column}{0.618\linewidth}
-			\begin{figure}
-
-				\only<1>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019_p_distribution.pdf}}
-				\only<2>{\includegraphics[width=\columnwidth]{../../figures/Appenzeller-Herzog_2019.pdf}}
-				\only<3>{\includegraphics[width=\columnwidth]{../../figures/Bos_2018.pdf}}
-				\only<4>{\includegraphics[width=\columnwidth]{../../figures/Brouwer_2019.pdf}}
-				\only<5>{\includegraphics[width=\columnwidth]{../../figures/Chou_2004.pdf}}
-				\only<6>{\includegraphics[width=\columnwidth]{../../figures/Donners_2021.pdf}}
-				\only<7>{\includegraphics[width=\columnwidth]{../../figures/Jeyaraman_2020.pdf}}
-			\end{figure}
-		\end{column}
-		\begin{column}{0.382\linewidth}
+		\begin{column}{0.32\linewidth}
 			\begin{itemize}
-				\item<1-> We generated LLM inclusion scores for each document in a set of reviews in the Synergy dataset
-				\item<2-> We compare what happens to screening in descending order of this ranking to 100 active learning runs with a default configuration of SVM with Tfidf
-				\item<3-> Initial results are not very promising
-				\item<5->But LLMs can do well where traditional methods struggle
-				\item<7->Sometimes they are no use at all
+				\item<1->Next, we calculate scores for our stopping criteria with a range of different recall targets
+				\item<2->Stopping when $p<\alpha$ gives a recall > target in the $\alpha^{th}$ percentile, in 97/100 settings
+				\item<3->In most settings, total work savings were greater for SVMs 
 			\end{itemize}
 		\end{column}
-	\end{columns}
+		\begin{column}{0.68\linewidth}
+			\only<2>{
+			\begin{table}
+				\scriptsize
+				\input{../../tables/recall.tex}
 
-\end{frame}
-
-\begin{frame}{All results}
-	\begin{columns}
-	\begin{column}{0.618\linewidth}
-		\begin{figure}
-
-			\only<1->{\includegraphics[width=\columnwidth]{../../figures/llm_svm_roc.pdf}}
-
-		\end{figure}
-	\end{column}
-	\begin{column}{0.382\linewidth}
-		\begin{itemize}
-			\item<1-> LLMs and basic active learning pipelines seem to have different weakneses
-			\item<2-> Combining both could improve general performance
-			\item<3-> LLMs seem most useful for smaller datasets (where active learning has little time to learn)
-		\end{itemize}
+			\end{table}
+		}
+			\only<3>{
+	\begin{table}
+		\scriptsize
+		\input{../../tables/wsst.tex}
+						\caption{absolute work savings across datasets (N records=169,288)}
+	\end{table}
+}
 		\end{column}
@@
 	\item<1-> Prompting strategies (inclusion criteria)
-	\item<2-> Bigger/different models
-	\item<3-> Combining LLMs with traditional approaches
-	\item<4-> Updating prompts based on user feedback
+	\item<2-> Combining LLMs with traditional approaches
+	\item<3-> Updating prompts based on user feedback
+	\item<4-> Better baselines (finetuning BERT)
 \end{itemize}
+
+\only<5->{Note that although the LLM implementation is simplistic, so is the comparator}
 
@@
 				\item<1-> LLMs are neither a quick fix or a silver bullet
-				\item<2-> Evaluation is vital
-				\item<3-> We can't forget the need for appropriate stopping criteria \cite{callaghan_statistical_2020}
-				\item<4->\url{mcallaghan.github.io/buscar-app}
+				\item<2-> They do not have magic properties that fix problems of traditional ML, they are simply likely to make them worse
+				\item<3-> We need to be on our guard!
+				\item<4-> Do not forget stopping criteria!
+				\item<5-> But they have some advantages, such as the fact that we only need to calculate scores once!
+				%\item<4->\url{mcallaghan.github.io/buscar-app}
 			\end{itemize}
@@
 			\begin{figure}
-				\only<3>{\includegraphics[width=\columnwidth]{../../figures/stopping.pdf}}
+				\only<4>{\includegraphics[width=\columnwidth]{../../figures/stopping.pdf}}
 			\end{figure}
@@
 
-\begin{frame}
-	Thanks!
-
-	\bigskip
-
-	\hrule
-
-	\bigskip
-
-	\url{callaghan@mcc-berlin.net}
-
-	\medskip
-
-	Twitter: @MaxCallaghan5
-
-	\bigskip
-
-	\begin{figure}
-	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
-	\end{figure}
-\end{frame}
+%	
+%\begin{frame}
+%	Thanks!
+%	
+%	\bigskip
+%	
+%	\hrule
+%	
+%	\bigskip
+%	
+%	\url{callaghan@mcc-berlin.net}
+%	 
+%	\medskip
+%	
+%	Twitter: @MaxCallaghan5
+%	
+%	\bigskip
+%	
+%	\begin{figure}
+%	\includegraphics[width=0.1\columnwidth]{images/wellcome-logo-black.eps}
+%	\end{figure}
+%\end{frame}
+%
 

### 2024-11-04 LLMs/manuscript/main.tex
Updates from Overleaf

@@
 		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use robust stopping criteria to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
-		We find that that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can save on average [x] percentage points of additional work compared to a baseline model.
-		Across the whole database of [x] systematic review datasets, which would have required [x] hours to screen manually, LLM-assisted screening can reduce the time required to [x] hours, with the next best model reducing the time required to [y] hours, while still ensuring a 95\% chance of missing no more than 5\% of relevant studies.
+		We find that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can produce ranked lists of documents that are superior to those generated by the baseline.
+		However, LLMs do not seem to be a silver bullet which would substantially reduce the amount of work required to complete screening.
+Across a variety of recall targets and confidence levels, work savings were generally higher with the SVM baseline, though possible work savings with LLMS were higher in low recall or low confidence settings.
 	\end{abstract}
@@
 	However, the paper proposes a strategy to save work using the model that is not able to accurately quantify or mitigate the risk of missing studies.
-Documents with an inclusion score greater than a threshold paramater $\theta$ are screened by hand, while all remaining documents are automatically excluded.
-The methods proposed to set the threshold paramater (either extrapolating from other reviews, or by setting the paramater such that it achieves the desired recall level on a set of predefined relevant documents) are \textit{not} able to reliably ensure that a desired level of recall is met.
+Documents with an inclusion score greater than a threshold parameter $\theta$ are screened by hand, while all remaining documents are automatically excluded.
+The methods proposed to set the threshold parameter (either extrapolating from other reviews, or by setting the parameter such that it achieves the desired recall level on a set of predefined relevant documents) are \textit{not} able to reliably ensure that a desired level of recall is met.
 
@@
 This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.
+
+    Similarly, a paper by Xia and co-authors \cite{xia_llmscreen_2024} highlights the lack of commonly agreed upon stopping criteria as a limitation of current approaches to ml-assisted screening, and proposes LLM-based screening as a solution.
 
@@
 
-	\bibliography{../stopping-criteria}
+	\bibliography{LLMs/stopping-criteria}
 	%\bibliography{../LLMs}

### 2024-11-04 LLMs/manuscript/main.tex

@@
 	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of this process, in order to reduce the labour required to complete a systematic review.
-However, much of this literature demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}.
+However, much of this literature introduces new tools, or demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}  \cite{van_de_schoot_open_2021, rathbone_faster_2015, yang_goldilocks_2022}.
 
@@
 
-	Early attempts to automate systematic review screening used simple models like logistic regression and support vector machines.
-Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs).
+	Early attempts to automate systematic review screening used simple models like logistic regression and support vector machines \cite{cohen_reducing_2006}.
+Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs) \cite{oami_performance_2024, wang_zero-shot_2024, luo_evaluating_2024, xia_llmscreen_2024}.
 However, not enough attention has been paid to comparing LLM-based approaches with the much simpler (and computationally less expensive) alternatives.
@@
     Similarly, a paper by Xia and co-authors \cite{xia_llmscreen_2024} highlights the lack of commonly agreed upon stopping criteria as a limitation of current approaches to ml-assisted screening, and proposes LLM-based screening as a solution.
+However, replacing one type of model with another does not solve the problem of how to quantify the risk of model error negatively affecting the study through missed reviews.
+Large language models do not possess properties that mean that we do not need to think carefully about how they are evaluated and used responsibly.
 

### 2024-11-07 LLMs/manuscript/main.tex
Updates from Overleaf

@@
 \usepackage{setspace}
-\usepackage{natbib}
+\usepackage[maxbibnames=5]{biblatex}
+%\usepackage{natbib}
 \usepackage{hyperref}
@@
 \usepackage[margin=2.5cm]{geometry}
+
+\addbibresource{LLMs/stopping-criteria.bib}
 
@@
 		For systematic review screening, responsible automation requires that the risks of missing relevant studies (which are inherent when the process is automated) can be accurately quantified and managed.
-		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use robust stopping criteria to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
-		We find that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can produce ranked lists of documents that are superior to those generated by the baseline.
+		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use a robust stopping criterion to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
+		We find that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can produce ranked lists of documents that are superior to those generated by our baseline - a standard active learning approach using a support vector machine.
 		However, LLMs do not seem to be a silver bullet which would substantially reduce the amount of work required to complete screening.
-Across a variety of recall targets and confidence levels, work savings were generally higher with the SVM baseline, though possible work savings with LLMS were higher in low recall or low confidence settings.
+Across a variety of recall targets and confidence levels, work savings were generally higher with the SVM baseline, though possible work savings with LLMs were higher in low recall or low confidence settings.
 	\end{abstract}
@@
 
-	Systematic Reviews are important tools to synthesise findings across multiple studies, enabling decision-makers to make decisions that are informed by the best available evidence. 
-	A key part of the process of systematic reviews involves finding studies using a transparent process that aims to be as comprehensive as possible. 
+	Rigorous systematic reviews methodologies are critical to synthesise findings across multiple studies, enabling decisions to be informed by the best available evidence. 
+	A key part of the systematic review process involves finding studies using a transparent process that aims to be as comprehensive as possible. 
 	Usually, bibliographic records are searched for using structured queries across multiple databases \cite{lefebvre_chapter_2023}.
 	This results in a list of records that may or not be relevant to the review.
-The titles and abstracts of these records are then examined by hand - often by two independent reviewers - and it is judged whether they should be included, which means they are read again at the full text level, or excluded.
-	This is a laborious and repetitive process - known as screening - and in times of increasing numbers of publications, this becomes more and more challenging.
+The titles and abstracts of these records are then examined by hand - often by two independent reviewers - and it is judged whether they should be included, which means they are read again at the full text level, or excluded, based on a transparent set of criteria demarcating the study boundaries.
+	This is a laborious and repetitive process - known as screening - and in times of increasing numbers of publications, this becomes more and more challenging \cite{bornmann_growth_2015}.
 
-	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of this process, in order to reduce the labour required to complete a systematic review.
+	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of the screening process, in order to reduce the labour required to complete a systematic review.
 However, much of this literature introduces new tools, or demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}  \cite{van_de_schoot_open_2021, rathbone_faster_2015, yang_goldilocks_2022}.
+This is insufficient.
+Evaluations in realistic settings are necessary for digital tools for the automation of systematic reviews to be used responsibly.
 
-	This is reflected in the fact that the guidelines for conducting systematic reviews published by the Cochrane colloquium \cite{lefebvre_chapter_2023} do not recommend that machine learning is used to save labour during screening.
+	This is reflected in the fact that the guidelines for conducting systematic reviews published by the Cochrane collaboration \cite{lefebvre_chapter_2023} do not recommend that machine learning is used to save labour during screening.
 This is because it remains uncertain how the use of automation may result in relevant studies being missed.
+This uncertainty will remain until the community reaches a consensus on what an adequate stopping criterion would look like, and what levels of confidence would be adequate to secure methodological standards.
 
-	Early attempts to automate systematic review screening used simple models like logistic regression and support vector machines \cite{cohen_reducing_2006}.
+	Early attempts to accelerate systematic review screening used simple models like logistic regression and support vector machines \cite{cohen_reducing_2006} in an active learning, or prioritised screening, setting.
+In such a setting, a subset of documents are screened by hand, and these documents are used to train a model to predict the relevance of remaining unscreened documents.
+A batch of those records with the highest predictions are then screened by hand, before the model is retrained using the expanded training dataset.
+This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.
+
  Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs) \cite{oami_performance_2024, wang_zero-shot_2024, luo_evaluating_2024, xia_llmscreen_2024}.
@@
 
-	Although the LLM-based procedure is compared to a fine-tuned BERT-like model, the details of this are not clear, and it appears that there is no comparison to the most common mode of machine learning assisted screening, known as machine learning prioritised screening.
-With ML-prioritised screening, a subset of documents are screened by hand, and these documents are used to train a model to predict the relevance of remaining unscreened documents.
-A batch of those records with the highest predictions are then screened by hand, before the model is retrained using the expanded training dataset.
-This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.
+	Although the LLM-based procedure is compared to a fine-tuned BERT-like model, the details of this are not clear, and it appears that there is no comparison to the most common mode of machine learning assisted screening, prioritised screening.
 
@@
 	In this paper, we adapt the probabilistic zero-shot screening process using LLMs to a ML-prioritised screening setting, and show that - when combined with statistically valid stopping criterion - it is able to reliably achieve high levels of recall while saving work.
-Secondly, we compare LLM-assisted approaches to simpler methods available, both in terms of the work saved, and in terms of computational requirements.
+Secondly, we compare LLM-assisted approaches to simpler methods available.
 
@@
 We simulate this process using the Synergy dataset of completed reviews \cite{de_bruin_synergy_2023}, and report on the work that would have been saved, and the actual level of recall that would have been achieved, given a range of recall targets and uncertainty levels.
+We do this using small (7/8 billion parameters) and medium (70 billion parameters) sized models from different generations of the Llama family of models \cite{touvron_llama_2023, dubey_llama_2024}.
+Llama models are open-weights, meaning they can freely be used and re-used, as opposed to proprietary models like GPT4.
 
@@
 	We implement our active learning pipeline as follows for each model and each review dataset, where each dataset has $N$ unique documents.
-Screening here, in a simulation context, refers to revealing the screening decision that was made when the dataset was created.
+Screening, in this simulation context, refers to revealing the screening decision that was made when the dataset was created.
 
@@
 
-	Overall, across the X datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}).
-Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer paramaters.
+	Overall, across the 25 systematic review datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}).
+Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer parameters.
 This difference was more pronounced for Llama2 than for Llama3.1.
@@
 
-	These results add to previous evidence that it is possible to \textit{safely} use prioritised screening approaches that can guarantee a given level of recall with a given level of confidence \cite{callaghan_statistical_2020}.
+	These results add to previous evidence that it is possible to \textit{safely} use prioritised screening approaches that, when combined with a robust stopping criterion, can guarantee a given level of recall with a given level of confidence \cite{callaghan_statistical_2020}.
 Across a wide range of datasets, as well as a wide range of different models, the stopping criteria ensured that the targeted level of recall was missed no more than expected given the confidence level.
@@
 
-	Simplified version, no inclusion criteria - but also simplified SVM setting, no parameter tuning, large batches, no finetuned.
+    Considering the environmental impact of ever larger models \cite{luccioni_power_2024}, these results merit a dose of caution before the evidence synthesis community rushes to adopt LLMs where simpler models may perform just as well.
+However, this is not to say that these approaches do not hold promise.
 
-	Possibility to combine
+    The results shown here are the product of a very simple implementation of LLM-assisted screening.
+That they were so impressive with a single zero-shot prediction for each document, using a very simple prompt that contains no additional information about the inclusion and exclusion criteria, hints that the use of LLMs may one day be able to generate substantial work savings.
+Future research may investigate the effects of using inclusion and exclusion criteria in the prompt, of more extensive "prompt tuning", and incorporating human labels into the prompt via in-context learning \cite{dong_survey_2024}.
 
-	Scientific treatment of the costs and benefits of different options.
+    Ultimately though, the utility of prioritisation improvements achieved by more complex models must be balanced against the financial and environmental costs of their deployment, as compared to simpler versions.
+It must be said also, that the baseline comparison - which consisted of few training iterations, a simple model, and no hyperparameter tuning - was also simplistic.
 
-	\bibliographystyle{unsrt}
+    Whatever approach is used, the question of how to mitigate the risks of automation to the quality of the study must be paramount.
+This is a question of statistics, not of machine learning engineering.
+Currently, prioritised screening with robust stopping criteria represents the best available method to manage the risk of missing studies.
+The stopping criteria we use has the benefit of ensuring that, should the model used perform less well than expected (but not worse than chance), it is only work savings that will be impacted, while the desired recall level is secured with the desired confidence level.
 
-	\bibliography{LLMs/stopping-criteria}
+	%\bibliographystyle{unsrt}
+
+	%\bibliography{LLMs/stopping-criteria}
 	%\bibliography{../LLMs}
+
+        \printbibliography
 

### 2024-11-08 LLMs/manuscript/main.tex
Updates from Overleaf

@@
 
-\title{Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?}
+\title{Can Zero-shot Large Language Models save additional work in machine learning prioritised screening for systematic reviews?}
 \author[1,2]{Max Callaghan}
+\author[1,2]{Jan Minx}
 
@@
 	\begin{abstract}
-		Zero shot learning using LLMs has received widespread attention in recent years, including in the automation of systematic review screening. 
+		Zero shot learning using Large Language Models (LLMs) has received widespread attention in recent years, including in the automation of systematic review screening. 
 		However, there has been insufficient research on robust processes to automate screening responsibly. 
@@
 Table \ref{tab:wssp} shows the mean and standard deviation of the work savings achieved for each combination of model, recall target, and confidence level.
-For low recall targets and low levels of confidence, large work savings are possible (>50\%).
+For low recall targets and low levels of confidence, large work savings are possible (\textgreater50\%).
 Increasing either the recall target or the level of confidence reduces the work savings that can be achieved.
@@
 
+    \section*{Code Availability}
+
+    The code required to reproduce the results of this study is made availabel at \url{https://github.com/mcallaghan/ml-screening-evaluation}
+
 	%\bibliographystyle{unsrt}
@@
 
+        \subsubsection*{Llama3.1}
+
+        \begin{verbatim}
+    <|begin_of_text|><|start_header_id|>system<|end_header_id|>
+    You are a systematic review helper tasked with finding out whether a study is relevant to the review $t
+
+    Answer 'yes' if the study is relevant, or 'no' if not
+    <|eot_id|><|start_header_id|>user<|end_header_id|>
+
+    Study: $s
+
+    Should the study be included? Answer yes or no. <|eot_id|>
+
+    <|start_header_id|>assistant<|end_header_id|>
+
+        \end{verbatim}
+
+        \subsubsection*{Llama2}
+        \begin{verbatim}
+    <s>[INST] <<SYS>>
+    You are a systematic review helper tasked with finding out 
+    whether a study is relevant to the review $t
+
+    Answer 'yes' if the study is relevant, or 'no' if not
+    <</SYS>>
+
+    Study: $s
+
+    Should the study be included? Answer yes or no. [/INST] 
+        \end{verbatim}
+
 \end{document}
